Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-08-18 2_SSL_NE/article/article_main.tex
added edits

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies, Anadon2016}. Among these drivers, the role of external knowledge and technology spillovers in research and development of energy technologies remains insufficiently studied \cite{Stephan2021}. While the exact definition of spillovers in the literature depends on the context \cite{Nemet2012, kolesnikov2020novel}, we specifically follow the approach of two previous studies \cite{Stephan2021, kolesnikov2020novel} by considering technology spillovers as the application of external knowledge in a technology where knowledge is considered external  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network beyond a particular technology may be an important factor in understanding \cite{Pichler2020, iea2020cleanenergy} and shaping \cite{Clark2016, Stephan2021} the future evolution of technologies.

>>>>

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies, Anadon2016}. Among these drivers, the role of external knowledge and technology spillovers in research and development of energy technologies remains insufficiently studied \cite{Stephan2021}. While the exact definition of spillovers in the literature depends on the context \cite{Nemet2012, kolesnikov2021spillovers}, we specifically follow the approach of two previous studies \cite{Stephan2021, kolesnikov2021spillovers} by considering technology spillovers as the application of external knowledge in a technology where knowledge is considered external  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network beyond a particular technology may be an important factor in understanding \cite{Pichler2020, iea2020cleanenergy} and shaping \cite{Clark2016, Stephan2021} the future evolution of technologies.

### 2023-08-18 2_SSL_NE/article/article_supplementary.tex
added edits

% cross-referencing
\input{2_SSL_NE/article/auxiliary/preamble/1_preamble_functionality_crossreference}

@@

Due to these limitations, any recommendations on future improvements in LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2020novel,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not yet been studied in the context of lighting.

@@

\begin{figure}[h!]
    \includegraphics[width=15cm]{./figures/sapphire_prices.pdf}
    \includegraphics[width=14.5cm]{./figures/wafer_size.pdf}
	\caption{\textbf{Historical trends in sapphire wafer size and price.} Top panel: Historical data for sapphire substrate prices of different surface properties and diameters, 2004-2020. Shown is data for non-patterned (polished) surface (NPS) and patterned surface substrates (PSS). Monocrystal denotes the Russian manufacturer of the same name. Dashed lines are projections made in 2015. Sources: (Monocrystal 2020)(Yole Developpement 2011)(Yole Developpement 2015). Bottom panel: Prevalence of sapphire wafer size used in LED manufacturing in 2008-2020. Hatched bars are projections. Sources: \cite{veeco2013}\cite{Scholand2012}\cite{yole2015sapphire}}
	\label{fig:wafers}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\end{figure}

\subsection{Non-Phosphor-Related LED Innovations}

@@

Importantly, the YAG phosphor does not exhibit the spectral properties desirable in general illumination applications (see Figure 6 in the main article). An early solution to this problem, which was first discovered in the late 1967 \cite{holloway1969optical} and suggested for LEDs by the same team at Nichia in 1996 \cite{bando1998development}\cite{shimizu1999light}, was to use the gadolinium-doped red-shifted YAG phosphor (YGAG). Used in combination with red-emitting sulfide phosphors, by 2003 it helped bring to the market the first generation of warm white light LED products, e.g., those produced by Lumileds \cite{Mueller2003}. However, sulfide phosphors led to accelerated deterioration of sensitive parts of LED devices and became less efficient as operating temperatures increased. New chemically stable and non-toxic red phosphors were needed.

@@

\begin{figure}[H]
	\centering
    \includegraphics[width=0.9\textwidth]{./figures/phosphor_spectrum-comparison.pdf}
	\caption{\textbf{Spectral data and additional consumer experience metrics for the earliest identified representative white LED products with published spectral data.} Corresponding LED products that used phosphor innovations listed in Table 1 in the main article are each indicated by the phosphor label and spectral data publication year. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Figure 6 in the main article. The desirable direction of improvements towards higher LER at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the bottom panels using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: spectral data for the first identified LED products that used phosphor innovations listed in Table 1 in the main article. The luminosity function \cite{cie-term-lumeff}, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable, as the photons of the corresponding energy are lost to the human eye and count towards the spectral loss channel. Plot legends indicate the years of publication of the spectral data and phosphor mixtures used in corresponding LED devices, with the following designations for additional phosphor mix components: ? -  other parts of phosphor mixture not disclosed, G$^*$ - \ch{CaSrS:Eu^{2+}}, Y$^\star$ - $\beta-$SiAlON:Eu$^{2+}$, G$^\dagger$ - \ch{Lu_3Al_5O_{12}:Ce^{3+}}, O$^\ddagger$ - \ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Source (top panel): own elaboration based on spectral data. Arb. Units - Arbitrary Units, defined as the ratio of radiation intensity at a wavelength compared to the highest point in the spectrum. Sources: top panel - own elaboration based on spectral data; bottom two panels: adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2003}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_calibration.pdf}
	\caption{\textbf{Cost structure comparison between our cost model, industry data and DOE projections.} Top panel: manufacturing cost calculations (2008-2016) and projections (2018, 2020) previously published by the U.S. DOE based on industry data and LEDCOM model. Bottom panel: manufacturing cost structure produced by our cost model for 2002, 2012 and 2020. Hatched bars are DOE projections. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \caption{List of sources for \cref{fgr:subeff} grouped by LED sub-efficiencies.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Panel (Sub-Eff.)} & \textit{References} \\
    \hline
    Panel A1 (FVE) & \cite{nichia2001data}\cite{lumi2003data}\cite{gen2005data}\cite{candlepwr2005data}\cite{lumi2006data}\cite{lumi2007data}\cite{nichia2008data}\cite{lumi2008data}\cite{osram2008data}\cite{jeong2011high}\cite{osram2012data}\cite{osram2013data}\cite{osram2014data} \newline \cite{lumi2016data_1}\cite{lumi2016data_2}\cite{epistar2017data}\cite{osram2017data_1}\cite{osram2017data_2}\cite{samsung2017data}\cite{samsung2018data}\cite{osram2018data}\cite{epistar2018data}\cite{lumi2019data} \\
    \hline
    Panel A2 (Droop) & Data calculated from luminous intensity curves of respective device datasheets: \cite{datasheet_osram_topled}\cite{osram2008data}\cite{osram2008gdplus}\cite{osram2018csp}\cite{datasheet_lumileds_lux1}\cite{lumi2008data}\cite{lumi2016data_1}\cite{lumi2016data_2}\cite{samsung2018data} \\
    \hline
    Panel B1 (IQE) & Data extracted from respective device datasheets and scientific publications (for a complete list, compare \cite{zenodo_weinold_led_history}). Additional data was extracted from government reports: \newline
\cite{doe_ssl_multiyear_2006}\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2009}\cite{doe_ssl_multiyear_2010}\cite{doe_ssl_multiyear_2011}\cite{doe_ssl_multiyear_2012}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_multiyear_2014}\cite{doe_ssl_rnd_2015}\cite{doe_ssl_rnd_2016} \\
    \hline
    Panel B2 (LEE) & \cite{lee2005analysis}\cite{krames2007status}\cite{Jang2004}\cite{Horng2013}\cite{Liao2010}\cite{HungWenHuang2005}\cite{Leem2007}\cite{Huang2008}\cite{Wang2009}\cite{Huh2003}\cite{Horng2008}\cite{Gao2008}\cite{Chang2003}\cite{Zhou2012} \newline \cite{ChunJuTun2006}\cite{Hua2009}\cite{Matioli2010}
\cite{lee2005analysis}\cite{Zhu2015}\cite{Ding2015}\cite{Taki2019}\cite{Shchekin2006}\cite{Hu2016}\cite{Horng2010}\cite{Lin2016}\cite{Yue2018}\cite{Zhao2012}\cite{Zhu2015}\newline \cite{Ding2015}\cite{wierer2001high}\cite{Steigerwald2003}\cite{DaeSeobHan2006}\cite{Wang2006}\cite{Lee2007}\cite{Shen2007}\cite{Huang2006}\cite{Zhmakin2011} \\
    \hline
    \end{NiceTabularX}
\end{table}

>>>>

@@

Due to these limitations, any recommendations on future improvements in LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2021spillovers,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not yet been studied in the context of lighting.

@@

\begin{figure}[h!]
    \includegraphics[width=15cm]{./figures/sapphire_prices.pdf}
    \includegraphics[width=14.5cm]{./figures/wafer_size.pdf}
	\caption{\textbf{Historical trends in sapphire wafer size and price.} Top panel: Historical data for sapphire substrate prices of different surface properties and diameters, 2004-2020. Shown is data for non-patterned (polished) surface (NPS) and patterned surface substrates (PSS). Monocrystal denotes the Russian manufacturer of the same name. Dashed lines are projections made in 2015. Sources: \cite{monocrystal2020private}\cite{yole2011sapphire}\cite{yole2015sapphire}. Bottom panel: Prevalence of sapphire wafer size used in LED manufacturing in 2008-2020. Hatched bars are projections. Sources: \cite{veeco2013}\cite{Scholand2012}\cite{yole2015sapphire}}
	\label{fig:wafers}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\end{figure}

\clearpage
\subsection{Non-Phosphor-Related LED Innovations}

@@

Importantly, the YAG phosphor does not exhibit the spectral properties desirable in general illumination applications (see Figure 6 in the main article). An early solution to this problem, which was first discovered in the late 1967 \cite{holloway1969optical} and suggested for LEDs by the same team at Nichia in 1996 \cite{bando1998development}\cite{shimizu1999light}, was to use the gadolinium-doped red-shifted YAG phosphor (YGAG). Used in combination with red-emitting sulfide phosphors, by 2003 it helped bring to the market the first generation of warm white light LED products, e.g., those produced by Lumileds \cite{Mueller2002}. However, sulfide phosphors led to accelerated deterioration of sensitive parts of LED devices and became less efficient as operating temperatures increased. New chemically stable and non-toxic red phosphors were needed.

@@

\begin{figure}[H]
	\centering
    \includegraphics[width=0.9\textwidth]{./figures/phosphor_spectrum-comparison.pdf}
	\caption{\textbf{Spectral data and additional consumer experience metrics for the earliest identified representative white LED products with published spectral data.} Corresponding LED products that used phosphor innovations listed in Table 1 in the main article are each indicated by the phosphor label and spectral data publication year. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Figure 6 in the main article. The desirable direction of improvements towards higher LER at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the bottom panels using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: spectral data for the first identified LED products that used phosphor innovations listed in Table 1 in the main article. The luminosity function \cite{cie-term-lumeff}, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable, as the photons of the corresponding energy are lost to the human eye and count towards the spectral loss channel. Plot legends indicate the years of publication of the spectral data and phosphor mixtures used in corresponding LED devices, with the following designations for additional phosphor mix components: ? -  other parts of phosphor mixture not disclosed, G$^*$ - \ch{CaSrS:Eu^{2+}}, Y$^\star$ - $\beta-$SiAlON:Eu$^{2+}$, G$^\dagger$ - \ch{Lu_3Al_5O_{12}:Ce^{3+}}, O$^\ddagger$ - \ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Source (top panel): own elaboration based on spectral data. Arb. Units - Arbitrary Units, defined as the ratio of radiation intensity at a wavelength compared to the highest point in the spectrum. Sources: top panel - own elaboration based on spectral data; bottom two panels: adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_calibration.pdf}
	\caption{\textbf{Cost structure comparison between our cost model, industry data and DOE projections.} Top panel: manufacturing cost calculations (2008-2016) and projections (2018, 2020) previously published by the U.S. DOE based on industry data and LEDCOM model. Bottom panel: manufacturing cost structure produced by our cost model for 2002, 2012 and 2020. Hatched bars are DOE projections. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel_details}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \caption{List of sources for \cref{fgr:subeff} grouped by LED sub-efficiencies.}
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Panel (Sub-Eff.)} & \textit{References} \\
    \hline
    Panel A1 (FVE) & \cite{nichia2001data}\cite{lumi2003lumi}\cite{gen2005data}\cite{candlepwr2005data}\cite{lumi2006data}\cite{lumi2007data}\cite{nichia2008data}\cite{lumi2008data}\cite{osram2008data}\cite{jeong2011high}\cite{osram2012data}\cite{osram2013data}\cite{osram2014data} \newline \cite{lumi2016data_1}\cite{lumi2016data_2}\cite{epistar2017data}\cite{osram2017data_1}\cite{osram2017data_2}\cite{samsung2017data}\cite{samsung2018data}\cite{osram2018data}\cite{epistar2018data}\cite{lumi2019data} \\
    \hline
    Panel A2 (Droop) & Data calculated from luminous intensity curves of respective device datasheets: \cite{datasheet_osram_topled}\cite{osram2008data}\cite{osram2008gdplus}\cite{osram2018csp}\cite{datasheet_lumileds_lux1}\cite{lumi2008data}\cite{lumi2016data_1}\cite{lumi2016data_2}\cite{samsung2018data} \\
    \hline
    Panel B1 (IQE) & Data extracted from respective device datasheets and scientific publications (for a complete list, compare \cite{zenodo_weinold_led_history}). Additional data was extracted from government reports: \newline
\cite{doe_ssl_multiyear_2006}\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2009}\cite{doe_ssl_multiyear_2010}\cite{doe_ssl_multiyear_2011}\cite{doe_ssl_multiyear_2012}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_multiyear_2014}\cite{doe_ssl_rnd_2015}\cite{doe_ssl_rnd_2016} \\
    \hline
    Panel B2 (LEE) & \cite{lee2005analysis}\cite{krames2007status}\cite{Jang2004}\cite{Horng2013}\cite{Liao2010}\cite{HungWenHuang2005}\cite{Leem2007}\cite{Huang2008}\cite{Wang2009}\cite{Huh2003}\cite{Horng2008}\cite{Gao2008}\cite{Chang2003}\cite{Zhou2012} \newline \cite{ChunJuTun2006}\cite{Hua2009}\cite{Matioli2010}
\cite{lee2005analysis}\cite{Zhu2015}\cite{Ding2015}\cite{Taki2019}\cite{Shchekin2006}\cite{Hu2016}\cite{Horng2010}\cite{Lin2016}\cite{Yue2018}\cite{Zhao2012}\cite{Zhu2015}\newline \cite{Ding2015}\cite{wierer2001high}\cite{Steigerwald2002}\cite{DaeSeobHan2006}\cite{Wang2006}\cite{Lee2007}\cite{Shen2007}\cite{Huang2006}\cite{Zhmakin2011} \\
    \hline
    \end{NiceTabularX}
\end{table}
